Towards a Mechanized Proof of Noninterference

### 2013-12-19 paper/paper.tex
some additions to the mechanizing proofs section

As mentioned above, we unfortunately were unable to complete our noninterference
proof in Coq by the time of this report.  Given more time, we are confident that
this could be completed however.  So far, we have proven the first few lemmas
described in Section~\ref{sec:proof} and defined relations for the operational
semantics and type system as presented in Section~\ref{sec:lang}.  This was much
more difficult than the short natural language proofs would suggest, due to a
combination of the stricter format required by Coq and the number of items that
are elided in the proofs because they were considered trivial or obvious.

We found that the process of mechanizing the lemmas that we proved was a useful
exercise in checking our understanding of our language, statements, and the work
of Simonet and Pottier which we based our approach off of.  We found ourselves
often surprised by the number of items that were elided by the proofs in the
paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
natural languae but turned out to be 40 lines of Coq code, not including all of
the supporting definitions and rules.

@@

We discovered a few small inaccuracies in \cite{InfoFlowML} by running into
issues when formalizing the proofs in Coq.  Listed below are the primary
corrections we were able to make.

@@

    As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
    structure for \langName from \origLang.  Specifically, we constructed
    values, answers, and expressions explicitly to not allow nesting of
    bracketed values.  This was not how the syntax figures in Simonet and
    Pottier's paper defined this, and as a result, we had trouble when our
    proofs depended on this explicit non-nesting property that was mentioned in
    the prose of their paper.  Realizing that our definitions did not disallow
    this by construction after running into issues in the proof, we were able to
    fix the definition producing what is in Figure~\ref{fig:syntax}.

@@

    Proving our own version of Lemma~\ref{lem:soundness} actually turned out to
    be instructive because it forced us to recognize that the original
    formulation in Simonet and Pottier's paper on \origLang wasn't exactly
    correct as stated.  Specifically, it is entirely possible that a step in the
    operational semantics does not translate to a step in the operational
    semantics under either projection, as in the case of the \ref{eq:lift-app}
    rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.  Luckily, we are
    fairly confident, given that we were still able to prove
    Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
    critical error in the original paper.
\end{description}

\todo{more?}

>>>>

As mentioned above, we unfortunately were unable to complete our noninterference
proof in Coq by the time of this report.  As of this writing, we have written definitions of a small-step operational semantics, a big-step operational semantics, typing judgments, and typing derivations for \langName.  We have proven a few lemmas from section~\ref{sec:proof} relating to the operations semantics (including soundness), but remain mired in complexity on the remainder of the mechanization effort.

In the process of mechanizing these lemmas, we empirically observed that the process of translating English into Coq can involve a nearly fifty-fold increase in the size of even the most basic proofs.  While this was partially a natural effect of writing proofs in Coq, we found that much of the complexity stemmed from handling cases elided in the English-Language proofs; and in these cases we found hidden subtleties which appear to have escaped even Simonet and Pottier.

@@

We discovered a few small inaccuracies in \cite{InfoFlowML} during the 
process of formalizing our proofs in Coq.  Listed below are the primary
corrections we were able to make.

While some inaccuracies did invalidate Simonet and Pottier's original
Lemmas, we universally found that a slightly different formulation was
provable and remained usable in future proofs.

@@

    As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
    structure for \langName ~than was seen in~ \origLang.  Specifically, we constructed
    values, answers, and expressions explicitly to not allow nesting of
    bracketed values.  The syntax defined by Simonet and Pottier did allow
    for this nesting, and authors dismissed is as a non-sensible
    construction.  Unfortunately, allowing such syntactic nesting 
    prevents the completeness Lemma from being proveable.
    Realizing that our definitions did not disallow
    this by construction after running into issues in the proof, we were able to
    fix the definition, producing what is in Figure~\ref{fig:syntax}.

@@

    Proving our own version of Lemma~\ref{lem:soundness} was uniquely
    instructive because it forced us to recognize that the original
    formulation in Simonet and Pottier's paper on \origLang ~wasn't 
    correct as stated.  Simonet and Pottier's Lemma asserts that 
    if $e \rightarrow e'$, then $\lift{e}_i \rightarrow \lift{e'}_i$.
    This does not hold in the case where $e$ is of the form $<v1 | v2> v$;
    In this case, the single reduction rule that applies is \ref{eq:lift-app},
    which does not step $\lift{e}_i \rightarrow \lift{e'}_i$ for any $i$ 
    (as $\lift{e}_i$ and $\lift{e'}_i$ are identical).  Luckily, we are
    fairly confident, given that we were still able to prove
    Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
    critical error in the original paper.
\end{description}
